### 2021-07-03 ch3bedEle.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

The second type of stochastic model prescribes rates (probabilities per unit time) to the erosion and deposition events of individual grains within a control volume to calculate the particle activity \citep{Einstein1950}.
These approaches aim at a complete statistical characterization of the bedload flux \citep{Furbish2012a,Fathel2015,Furbish2017,Heyman2016}, including probability distributions \citep{Ancey2006,Ancey2008}, spatial and temporal characteristics of its fluctuations \citep{Ancey2008, Heyman2014a, Roseberry2012,Dhont2018}, and the dependence of these statistical characteristics on the length and time scales over which they are measured or calculated \citep{Singh2009,Singh2012,Saletti2015,Ma2014}.
A recent surge in research activity has generated rapid progress and spawned many new inquiries in this subject. 
For example, \citet{Ancey2006} demonstrated that a constant erosion rate as originally proposed by \citet{Einstein1950} was insufficient to develop realistically large particle activity fluctuations, so they added a positive feedback between the particle activity and erosion rate they called ``collective entrainment'' \citep{Ancey2008, Heyman2013,Heyman2014,Ma2014, Lee2018}.
While they deemed this feedback necessary to model realistic activity fluctuations, the implications of this collective entrainment term on bed topography and particle activity changes has not been fully explored.

@@

We write the master equation for the probability flow using the forward Kolmogorov equation $\partial P(n,m;t)/\partial t = 
\sum_{n',m'} [R(n,m|n',m')P(n',m';t)-R(n',m'|n,m)P(n,m;t)]$ \citep{Cox1965, Gillespie1992, Ancey2008} as 
\begin{multline}
	\frac{\partial P}{\partial t}(n,m;t) =  
	\nu P(n-1,m;t) + [\lambda(m+1) + \mu(n-1)][1+\kappa(m+1)]P(n-1,m+1;t)\\  
	+ \sigma(n+1)[1-\kappa(m-1)]P(n+1,m-1;t) + \gamma(n+1) P(n+1,m;t) \\
	- 
	\{ \nu + \lambda+ \mu n (1+\kappa m) +  \sigma n ( 1- \kappa m) + \gamma n \}P(n,m;t).
	\label{eq:master}
\end{multline}
The joint probability distribution $P(n,m;t)$ solving this equation fully characterizes the statistics of $n$ and $m$ -- proxies for the bedload flux and local bed elevation.
The average entrainment and deposition rates $E$ and $D$ over all bed elevations are $E = \lambda +\mu \langle n \rangle$ and $D=\sigma \langle n \rangle$.
We anticipate that solutions of (\ref{eq:master}) will adjust from the initial conditions to a steady-state distribution $P_s(n,m)$ -- independent of time -- if the constant factors in the transition rates are representative of equilibrium conditions. Equilibrium requires $E=D$, meaning there is no net change in elevation, and $\nu = \gamma \langle n \rangle$, meaning mass is conserved in the control volume (inflow $=$ outflow).
This Master equation describes a two-species stochastic birth-death model \citep{Cox1965} of a type well-known in population ecology \citep{Pielou1977, Swift2002} and chemical physics \citep{Gardiner1983}.
In our context, the two populations are the moving and stationary grains in the volume.

@@

Unfortunately, equation (\ref{eq:master}) does not appear to admit an analytical solution unless $\kappa=0$ (but see \citet{Swift2002} for the generating function method which fails in this case).
The difficulty originates from the product terms between $n$ and $m$ representing the bed elevation dependence of collective entrainment and deposition rates.
In response to this difficulty, we resort to numerical methods and analytical approximations, simulating equation (\ref{eq:master}) with the Gillespie algorithm \citep{Gillespie1977, Gillespie1992, Gillespie2007} and solving it approximately with mean field and Fokker-Planck approaches \citep{Haken1983,Gardiner1983}. The simulation algorithm is described in \ref{sec:numerical}, and analytical approximations are described in \ref{sec:analytical}.

@@

In this paper, we have built on earlier works \citep{Ancey2008,Martin2014} to include the vertical dimension of bed elevation dynamics, study the interplay between bedload transport and bed elevation fluctuations, and investigate resting time distributions of sediment undergoing burial.
To our knowledge, this model is the first description of bedload transport and bed elevations as a coupled stochastic population model based on individual grains.
Numerical solutions and analytical approximations provided negative binomial distributions of bedload activity and normal distributions of bed elevations.
Although experiments under more natural conditions with segregation processes,  migrating bedforms, or sediment supply perturbations have shown particle activity distributions with heavier tails \citep{Dhont2018,Saletti2015a} and non-Gaussian bed elevations \citep{Singh2012,Aberle2006}, our results reproduce the key features of the most controlled bed elevation \citep{Wong2007,Martin2014} and bedload transport \citep{Heyman2016,Ancey2008} experiments in the literature.

@@

Like \citet{Martin2014}, we obtained heavy-tailed power-law resting times for sediment undergoing burial by treating bed elevation changes as an unbounded random walk with a mean reverting tendency.
This result suggests sediment burial can explain the heavy-tailed rests seen in field data \citep{Olinde2015,Bradley2017,Pretzlav2016}. 
Our resting time distributions show a divergent variance and possibily a divergent mean, since this occurs for $\alpha < 1$ \citep{Sornette2000} which is within range of our results.
Divergent mean resting time distributions present a paradox, since they imply all particles should eventually be immobile, violating the equilibrium transport assumption.
\citet{Voepel2013} demonstrated that a bounded random walk for bed elevations provides a power-law distribution that eventually transitions to a faster thin-tailed decay, allowing for power-law scaling like our result and \citet{Martin2014} without this divergent mean paradox. 
One resolution to this issue could come from a spatially distributed model with multiple cells. Neighboring locations might bound excessive local elevation changes through granular relaxations from gradients above the angle of repose. In this interpretation, divergent mean power law resting time distributions may be relics of single cell models for bed elevation changes. We should always expect a maximum depth to which the bed can degrade relative to neighboring locations; this could temper the power law tail without required the reflecting boundaries used by \citet{Voepel2013}.

Finally, we studied probability distribution functions and first and second moments of the particle activity and bed elevation, making novel conclusions about coordination between the statistical characteristics of these quantites which deserve experimental testing.
In the last decade, particle tracking experiments have emerged \citep{Lajeunesse2010,Roseberry2012,Martin2014, Fathel2015,Heyman2016,Liu2019}, that allow joint resolution of bed elevations and bedload transport. 
A suitably designed experiment could test our prediction that bed elevations regulate particle activity statistics, as essentially represented in figures \ref{fig:condmoms} and \ref{fig:colent}. 
However, we have left many other statistical characteristics of bedload transport for future studies. 
For example, the dependence of bedload transport \citep{Saletti2015a,Singh2009} and bed elevation statistics \citep{Aberle2006, Singh2009,Singh2012} on the spatial and temporal scales over which they are observed is an emerging research topic. Statistical quantities can either be monoscaling or multiscaling across the observation scale \citep{Sornette2000}, and we currently lack physical understanding and general conclusions about the scale dependence of particle activity and bed elevation signals.
The model we have presented shows statistical monoscaling for both quantites \citep[e.g.][]{Saletti2015a}, whereas other experiments indicate statistical multiscaling \citep{Aberle2006,Singh2009,Singh2012}. We consider this topic to go beyond the scope of the present work, and we have focused on statistical characteristics at the highest temporal resolutions, with no averaging over the observation scale.

>>>>

The second type of stochastic model prescribes rates (probabilities per unit time) to the erosion and deposition events of individual grains within a control volume to calculate the particle activity \citep{Einstein1950}.
These approaches aim at a complete statistical characterization of the bedload flux \citep{Furbish2012a,Fathel2015,Furbish2017,Heyman2016}, including probability distributions \citep{Ancey2006,Ancey2008}, spatial and temporal characteristics of its fluctuations \citep{Ancey2008, Heyman2014a, Roseberry2012,Dhont2018}, and the dependence of these statistical characteristics on the length and time scales over which they are measured or calculated \citep{Singh2009,Singh2012,Saletti2015}.
A recent surge in research activity has generated rapid progress and spawned many new inquiries in this subject. 
For example, \citet{Ancey2006} demonstrated that a constant erosion rate as originally proposed by \citet{Einstein1950} was insufficient to develop realistically large particle activity fluctuations, so they added a positive feedback between the particle activity and erosion rate they called ``collective entrainment'' \citep{Ancey2008, Heyman2013,Heyman2014,Ma2014, Lee2018}.
While they deemed this feedback necessary to model realistic activity fluctuations, the implications of this collective entrainment term on bed topography and particle activity changes has not been fully explored.

@@

We write the master equation for the probability flow using the forward Kolmogorov equation $\partial P(n,m;t)/\partial t = 
\sum_{n',m'} [R(n,m|n',m')P(n',m';t)-R(n',m'|n,m)P(n,m;t)]$ \citep{Cox1965, Gillespie1991, Ancey2008} as 
\begin{multline}
	\frac{\partial P}{\partial t}(n,m;t) =  
	\nu P(n-1,m;t) + [\lambda(m+1) + \mu(n-1)][1+\kappa(m+1)]P(n-1,m+1;t)\\  
	+ \sigma(n+1)[1-\kappa(m-1)]P(n+1,m-1;t) + \gamma(n+1) P(n+1,m;t) \\
	- 
	\{ \nu + \lambda+ \mu n (1+\kappa m) +  \sigma n ( 1- \kappa m) + \gamma n \}P(n,m;t).
	\label{eq:master}
\end{multline}
The joint probability distribution $P(n,m;t)$ solving this equation fully characterizes the statistics of $n$ and $m$ -- proxies for the bedload flux and local bed elevation.
The average entrainment and deposition rates $E$ and $D$ over all bed elevations are $E = \lambda +\mu \langle n \rangle$ and $D=\sigma \langle n \rangle$.
We anticipate that solutions of (\ref{eq:master}) will adjust from the initial conditions to a steady-state distribution $P_s(n,m)$ -- independent of time -- if the constant factors in the transition rates are representative of equilibrium conditions. Equilibrium requires $E=D$, meaning there is no net change in elevation, and $\nu = \gamma \langle n \rangle$, meaning mass is conserved in the control volume (inflow $=$ outflow).
This Master equation describes a two-species stochastic birth-death model \citep{Cox1965} of a type well-known in population ecology \citep{Pielou1977, Swift2002} and chemical physics \citep{Gardiner1983}.
In our context, the two populations are the moving and stationary grains in the volume.

@@

Unfortunately, equation (\ref{eq:master}) does not appear to admit an analytical solution unless $\kappa=0$ (but see \citet{Swift2002} for the generating function method which fails in this case).
The difficulty originates from the product terms between $n$ and $m$ representing the bed elevation dependence of collective entrainment and deposition rates.
In response to this difficulty, we resort to numerical methods and analytical approximations, simulating equation (\ref{eq:master}) with the Gillespie algorithm \citep{Gillespie1977, Gillespie1991, Gillespie2007} and solving it approximately with mean field and Fokker-Planck approaches \citep{Haken1978,Gardiner1983}. The simulation algorithm is described in \ref{sec:numerical}, and analytical approximations are described in \ref{sec:analytical}.

@@

In this paper, we have built on earlier works \citep{Ancey2008,Martin2014} to include the vertical dimension of bed elevation dynamics, study the interplay between bedload transport and bed elevation fluctuations, and investigate resting time distributions of sediment undergoing burial.
To our knowledge, this model is the first description of bedload transport and bed elevations as a coupled stochastic population model based on individual grains.
Numerical solutions and analytical approximations provided negative binomial distributions of bedload activity and normal distributions of bed elevations.
Although experiments under more natural conditions with segregation processes,  migrating bedforms, or sediment supply perturbations have shown particle activity distributions with heavier tails \citep{Dhont2018,Saletti2015} and non-Gaussian bed elevations \citep{Singh2012,Aberle2006}, our results reproduce the key features of the most controlled bed elevation \citep{Wong2007,Martin2014} and bedload transport \citep{Heyman2016,Ancey2008} experiments in the literature.

@@

Like \citet{Martin2014}, we obtained heavy-tailed power-law resting times for sediment undergoing burial by treating bed elevation changes as an unbounded random walk with a mean reverting tendency.
This result suggests sediment burial can explain the heavy-tailed rests seen in field data \citep{Olinde2015,Bradley2017,Pretzlav2016a}. 
Our resting time distributions show a divergent variance and possibily a divergent mean, since this occurs for $\alpha < 1$ \citep{Sornette2006} which is within range of our results.
Divergent mean resting time distributions present a paradox, since they imply all particles should eventually be immobile, violating the equilibrium transport assumption.
\citet{Voepel2013} demonstrated that a bounded random walk for bed elevations provides a power-law distribution that eventually transitions to a faster thin-tailed decay, allowing for power-law scaling like our result and \citet{Martin2014} without this divergent mean paradox. 
One resolution to this issue could come from a spatially distributed model with multiple cells. Neighboring locations might bound excessive local elevation changes through granular relaxations from gradients above the angle of repose. In this interpretation, divergent mean power law resting time distributions may be relics of single cell models for bed elevation changes. We should always expect a maximum depth to which the bed can degrade relative to neighboring locations; this could temper the power law tail without required the reflecting boundaries used by \citet{Voepel2013}.

Finally, we studied probability distribution functions and first and second moments of the particle activity and bed elevation, making novel conclusions about coordination between the statistical characteristics of these quantites which deserve experimental testing.
In the last decade, particle tracking experiments have emerged \citep{Lajeunesse2010,Roseberry2012,Martin2014, Fathel2015,Heyman2016,Liu2019}, that allow joint resolution of bed elevations and bedload transport. 
A suitably designed experiment could test our prediction that bed elevations regulate particle activity statistics, as essentially represented in figures \ref{fig:condmoms} and \ref{fig:colent}. 
However, we have left many other statistical characteristics of bedload transport for future studies. 
For example, the dependence of bedload transport \citep{Saletti2015,Singh2009} and bed elevation statistics \citep{Aberle2006, Singh2009,Singh2012} on the spatial and temporal scales over which they are observed is an emerging research topic. Statistical quantities can either be monoscaling or multiscaling across the observation scale \citep{Sornette2006}, and we currently lack physical understanding and general conclusions about the scale dependence of particle activity and bed elevation signals.
The model we have presented shows statistical monoscaling for both quantites \citep[e.g.][]{Saletti2015}, whereas other experiments indicate statistical multiscaling \citep{Aberle2006,Singh2009,Singh2012}. We consider this topic to go beyond the scope of the present work, and we have focused on statistical characteristics at the highest temporal resolutions, with no averaging over the observation scale.

### 2021-07-03 ch4downDiff.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

Nikora et al. explained that divergent resting times generate a sub-diffusive global range.
However, studies have demonstrated that divergent resting times can generate super-diffusion in asymmetric random walks \citep{Weeks1996,Weeks1998}, and both experiments \citep{Bradley2017,Bradley2010} and models \citep{Pelosi2016,Wu2019,Wu2019a} of bedload tracers undergoing burial have demonstrated global range super-diffusion.
While our results also show global range super-diffusion, they do not necessarily refute the Nikora et al. conclusion of sub-diffusion at long timescales.
We assumed sediment burial was a permanent condition which developed a non-diffusive geomorphic range.
In actuality, burial is a temporary condition, because bed scour can exhume buried sediment back into transport \citep{Wu2019a}, probably after heavy-tailed intervals \citep{Voepel2013,Martin2014,Pierce2020}.
We anticipate that a generalization of our model to include heavy-tailed timescales between burial and exhumation would develop four ranges of diffusion, where the long-time decay of the exhumation time distribution would dictate the geomorphic range diffusion characteristics as depicted in figure \ref{fig:var}.
If cumulative exhumation times decay faster than $T^{-1/2}$, as suggested by equilibrium transport models \citep{Voepel2013, Martin2014, Pierce2020a} and laboratory experiments \citep{Martin2014,Martin2012}, we expect a super-diffusive geomorphic range \citep{Weeks1998}.
However, if they decay slower than $T^{-1/2}$, as implicitly suggested by the data of \citet{Olinde2015}, we expect a genuinely sub-diffusive geomorphic range \citep{Weeks1998}, leaving Nikora et al. with the final word on long-time sub-diffusion.

>>>>

Nikora et al. explained that divergent resting times generate a sub-diffusive global range.
However, studies have demonstrated that divergent resting times can generate super-diffusion in asymmetric random walks \citep{Weeks1996,Weeks1998}, and both experiments \citep{Bradley2017,Bradley2010} and models \citep{Pelosi2016,Wu2019,Wu2019a} of bedload tracers undergoing burial have demonstrated global range super-diffusion.
While our results also show global range super-diffusion, they do not necessarily refute the Nikora et al. conclusion of sub-diffusion at long timescales.
We assumed sediment burial was a permanent condition which developed a non-diffusive geomorphic range.
In actuality, burial is a temporary condition, because bed scour can exhume buried sediment back into transport \citep{Wu2019a}, probably after heavy-tailed intervals \citep{Voepel2013,Martin2014}.
We anticipate that a generalization of our model to include heavy-tailed timescales between burial and exhumation would develop four ranges of diffusion, where the long-time decay of the exhumation time distribution would dictate the geomorphic range diffusion characteristics as depicted in figure \ref{fig:var}.
If cumulative exhumation times decay faster than $T^{-1/2}$, as suggested by equilibrium transport models \citep{Voepel2013, Martin2014, Pierce2020a} and laboratory experiments \citep{Martin2014,Martin2012}, we expect a super-diffusive geomorphic range \citep{Weeks1998}.
However, if they decay slower than $T^{-1/2}$, as implicitly suggested by the data of \citet{Olinde2015}, we expect a genuinely sub-diffusive geomorphic range \citep{Weeks1998}, leaving Nikora et al. with the final word on long-time sub-diffusion.

### 2021-07-03 ch5langevin.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

Fluid forces on a coarse particle in a viscous flow depend on the Reynolds number $\Re_p = d V/\nu$ defined by the particle size $d$, slip velocity $V$ between particle and fluid, and kinematic viscosity $\nu$.
These forces have been calculated analytically from the Navier-Stokes equations for vanishing $\Re_p$ and include acceleration, history, and velocity-dependent drag terms \citep{Hjelmfelt1966, Maxey1983, Auton1987}.
At realistic $\Re_p$ analytical results are limited, so it is standard practice to turn instead to empirical corrections on the small $\Re_p$ formulas \citep{Schmeeckle2007,Mei2006}.
A dominant contribution to the downstream drag force $F_D$ on nearly spherical particles at large $\Re_p$ can be written $F_D = \frac{\pi}{8}
\rho_f d^2 C_D(\Re_p) |V|V$, where $\rho_f$ is the fluid density, $d$ is the particle diameter, $C_D(\Re_p)$ is an empirical drag coefficient, and $V = U-u$ is the slip velocity between the fluid ($U$) and particle ($u$) velocities \citep{Coleman1967, Schmeeckle2007, Dwivedi2012}.
In the present model we set $C_D = \frac{24}{\Re_p}( 1 + 0.194 \Re_p^{0.631})$ \citep{Clift1978,Gonzalez2017} and we do not involve acceleration and history terms for simplicity, although we acknowledge their potential importance for coarse sediment transport \citep{Michaelides1997,Armenio2001,Dalche2015}.

@@

\subsection{Derivation of the bedload velocity distribution}
\label{sec:solution}
Hereafter we drop the prime on the most common streamwise restitution coefficient $\ve'$. With the truncation to two terms, equation \ref{eq:master} gives 
\be 0 = -\tilde{\Gamma}\partial_u P(u) + \tilde{D}\partial_u^2 P(u) -P(u) + \frac{1}{\ve} P\big(\frac{u}{\ve}\big),\label{eq:governer} \ee
which is now a non-local ordinary differential equation. Such equations have seen some attention in the mathematics literature, where they are called pantograph equations \citep{Hall1970,Zaidi2015,Bhalekar2017} for their relationship to a current collection device used on electric trains \citep{Ockendon1971}.
In the appendix we solve equation \ref{eq:governer} using Laplace transforms, providing
\begin{multline} P(u) = \frac{\theta(-u)}{K_+}\sum_{l=0}^\infty \frac{\ve^{-l}e^{\lambda_+\ve^{-l}u}}{\prod_{m=1}^l(-\tilde{D} \lambda_+^2 \ve^{-2m} + \tilde{\Gamma} \lambda_+ \ve^{-m} + 1) } 
	\\+ \frac{\theta(u)}{K_-}\sum_{l=0}^\infty \frac{\ve^{-l}e^{\lambda_-\ve^{-l}u}}{\prod_{m=1}^l(-d \lambda_-^2 \ve^{-2m} + \gamma \lambda_- \ve^{-m} + 1) } \label{eq:steadystate}. \end{multline}
The factors $\lambda_\pm$ are defined in the appendix; they are proportional to $\tilde{\Gamma}/\tilde{D}$. 
The normalization factors are $K_\pm$ are 
\be K_\pm = d(\lambda_+-\lambda_-)\prod_{l=1}^\infty (-d\lambda_\pm^2 \ve^{2l} + \gamma \lambda_\pm \ve^{l} + 1). \ee
Although this velocity distribution appears quite complicated, one can verify that this is a normalized probability distribution which has very simple limiting behaviors as the most common dissipation coefficient $\ve$ approaches fully elastic ($\ve=1$) and inelastic ($\ve = 0$) values.

@@

Many phenomena of landscape evolution depend on the velocities of individual grains.
The erosion rates of bedrock rivers are for example parameterized by the impact velocities of individual sediment grains against the bedrock surface \citep{Sklar2004,Tingian}, on which they have a sensitive nonlinear dependence.
It is not much of a stretch to imagine that bedrock erosion could be dominated by the largest impact velocities that originate from the extreme tail of the bedload velocity distribution. Here, we have shown that the weight of this tail depends sensitively on the dissipation characteristics of particle-bed collisions, suggesting that another feedback, that between dissipation and ... may be at play in the evolution of bedrock canyons.

>>>>

Fluid forces on a coarse particle in a viscous flow depend on the Reynolds number $\Re_p = d V/\nu$ defined by the particle size $d$, slip velocity $V$ between particle and fluid, and kinematic viscosity $\nu$.
These forces have been calculated analytically from the Navier-Stokes equations for vanishing $\Re_p$ and include acceleration, history, and velocity-dependent drag terms \citep{Hjelmfelt1966, Maxey1983, Auton1987}.
At realistic $\Re_p$ analytical results are limited, so it is standard practice to turn instead to empirical corrections on the small $\Re_p$ formulas \citep{Schmeeckle2007,Clift1978}.
A dominant contribution to the downstream drag force $F_D$ on nearly spherical particles at large $\Re_p$ can be written $F_D = \frac{\pi}{8}
\rho_f d^2 C_D(\Re_p) |V|V$, where $\rho_f$ is the fluid density, $d$ is the particle diameter, $C_D(\Re_p)$ is an empirical drag coefficient, and $V = U-u$ is the slip velocity between the fluid ($U$) and particle ($u$) velocities \citep{Coleman1967, Schmeeckle2007, Dwivedi2012}.
In the present model we set $C_D = \frac{24}{\Re_p}( 1 + 0.194 \Re_p^{0.631})$ \citep{Clift1978,Gonzalez2017} and we do not involve acceleration and history terms for simplicity, although we acknowledge their potential importance for coarse sediment transport \citep{Michaelides1997,Armenio2001}.

@@

\subsection{Derivation of the bedload velocity distribution}
\label{sec:solution}
Hereafter we drop the prime on the most common streamwise restitution coefficient $\ve'$. With the truncation to two terms, equation \ref{eq:master} gives 
\be 0 = -\tilde{\Gamma}\partial_u P(u) + \tilde{D}\partial_u^2 P(u) -P(u) + \frac{1}{\ve} P\big(\frac{u}{\ve}\big),\label{eq:governer} \ee
which is now a non-local ordinary differential equation. have seen some attention in the mathematics literature, where they are called pantograph equations \citep{Hall1989, Kim1998,Zaidi2015}.
In the appendix we solve equation \ref{eq:governer} using Laplace transforms, providing
\begin{multline} P(u) = \frac{\theta(-u)}{K_+}\sum_{l=0}^\infty \frac{\ve^{-l}e^{\lambda_+\ve^{-l}u}}{\prod_{m=1}^l(-\tilde{D} \lambda_+^2 \ve^{-2m} + \tilde{\Gamma} \lambda_+ \ve^{-m} + 1) } 
	\\+ \frac{\theta(u)}{K_-}\sum_{l=0}^\infty \frac{\ve^{-l}e^{\lambda_-\ve^{-l}u}}{\prod_{m=1}^l(-d \lambda_-^2 \ve^{-2m} + \gamma \lambda_- \ve^{-m} + 1) } \label{eq:steadystate}. \end{multline}
The factors $\lambda_\pm$ are defined in the appendix; they are proportional to $\tilde{\Gamma}/\tilde{D}$. 
The normalization factors are $K_\pm$ are 
\be K_\pm = d(\lambda_+-\lambda_-)\prod_{l=1}^\infty (-d\lambda_\pm^2 \ve^{2l} + \gamma \lambda_\pm \ve^{l} + 1). \ee
Although this velocity distribution appears quite complicated, one can verify that this is a normalized probability distribution which has very simple limiting behaviors as the mode dissipation constant $\ve$ approaches fully elastic ($\ve=1$) and inelastic ($\ve = 0$) values.

@@

Many phenomena of landscape evolution depend on the velocities of individual grains.
The erosion rates of bedrock rivers are for example parameterized by the impact velocities of individual sediment grains against the bedrock surface \citep{Sklar2004,Li2020,Turowski2021}, on which they have a sensitive dependence.
It is not much of a stretch to imagine that bedrock erosion could be dominated by the largest impact velocities that originate from the extreme tail of the bedload velocity distribution. Here, we have shown that the weight of this tail depends sensitively on the dissipation characteristics of particle-bed collisions, suggesting that another feedback, that between dissipation and ... may be at play in the evolution of bedrock canyons.

### 2021-07-03 ch6conc.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

\chapter{Summary and future work}
\label{ch:conc}
\section{The overall strategy}
\subsection{Langevin and master equations}
\subsection{Idealized noises and their combinations}

>>>>

\chapter{Summary and future work}
\label{ch:conc}
\section{The overall strategy}

This thesis has presented a review and four chapters which generalize and extend earlier work to describe sediment transport as a stochastic process. In the research, I characterized the movements of grains along a wider range of timescales than before, and I developed new a new formalism using these movements to predict the sediment flux.

\subsection{Langevin and master equations}

Probability

\subsection{Idealized noises and their combinations}

### 2021-07-03 pref.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
The thesis includes two published works, Chapters \ref{ch:bedEle} and \ref{ch:downdiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \citet{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication subsequent to the thesis submission.

>>>>

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
The thesis includes two published works, Chapters \ref{ch:bedEle} and \ref{ch:downdiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \textit{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication subsequent to the thesis submission.
